Formalizing Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2023-01-03 nakayama/paper.tex
Edits to description on Axiom of Choice

The notion of a module over a ring is an extension of the linear-algebraic
notion of a vector space over a field, ubiquitous in mathematics and its
applications. Less frequently referred to as the \emph{Krull-Azumaya theorem}
\footnote{Hideyuki Matsumura explains in his text \emph{Commutative Algebra}
  \cite{matsumura}: ``{This simple but important lemma is due to T.\ Nakayama,
  G.\ Azumaya, and W.\ Krull. Priority is obscure, and although it is usually
  called the Lemma of Nakayama, late Prof.\  Nakayama did not like the
  name.''}}
\,\cite{nagata}, Nakayama's Lemma describes one way that a finitely generated
module over an arbitrary commutative ring acts like a vector space over a
field. True to the convention that ``lemma'' often refers to a result serving
as a stepping stone toward another goal, Nakayama's Lemma is applied widely
throughout the field, and the result is typically introduced in a first
graduate course in commutative algebra \cite{atiyah-macdonald, matsumura,
eisenbud}.

@@

\subsection{Constructing the Formal Proof}
Beyond formalizing the relevant structures from higher algebra, we also
formally establish some basic theory. For instance, a \emph{unit} of a
commutative ring $R$ is an element $x \in R$ with a multiplicative inverse,
i.e., an element $x^{-1} \in R$ for which $x \cdot x^{-1}$ is the
multiplicative identity $1 \in R$.  In fact, it $x$ is a unit, it has a unique 
inverse.

We formally prove that an ideal $I$ of a commutative ring $R$ that contains a
unit $x$, then $I$ must be the trivial ideal, i.e., the entire ring.  The
informal logic is as follows:  Due to the absorption property of $I$,  
$x \cdot x^{-1} = 1 \in I$.  Hence for every element $r$ of $R$,
$r = r \cdot 1$ is also in $I$, i.e., $I=R$.

Every non-unit element of a commutative ring is contained in some maximal
ideal, a fact that relies on the Axiom of Choice.

The following standard informal
proof calls upon Zorn's lemma, which is equivalent to the Axiom of Choice, and
says that given a partially ordered set $S$, if every chain in $S$ has an upper
bound, then $S$ must have at least one maximal element.

\begin{quote}
    Set \(I_{1}\) to be the principal ideal generated by an element \(x\) of a
    commutative ring $R$, i.e. the smallest ideal containing the element $x$,
    which consists of all $R$-multiples of $x$.

    If \(I_{1}\) is not a maximal ideal, then there exists a strictly larger
    proper ideal \(I_{2}\) of $R$, i.e.,
    $x\in I_{1}\subsetneq I_{2} \subsetneq R$. Moreover, if  \(I_{2}\) is not
    maximal, then there exists a strictly larger proper ideal
    $I_{3}$ containing $I_2$.

    Continuing this process, if it terminates at some step, we have found a
    maximal ideal, and if not, one obtains an infinite chain of ideals
    containing \(x\).
    \[x\in I_{1}\subsetneq I_{2}\subsetneq I_{3}\subsetneq\cdots\subsetneq R\]

    It is straightforward to verify that the increasing union of all $I_{k}$,
    $k \geq 1$, is by definition an ideal of $R$, and certainly
    contains \(x\).  Hence by Zorn's lemma, since every ascending chain of
    ideals containing $x$ with respect to inclusion has its union as an upper
    bound, there exists a maximal ideal of \(R\) containing \(x\).
\end{quote}

>>>>

The notion of a module over a ring is an extension of the linear-algebraic
notion of a vector space over a field, ubiquitous in mathematics and its
applications. Less frequently referred to as the \emph{Krull-Azumaya theorem}
\footnote{Hideyuki Matsumura explains in his book \emph{Commutative Algebra}
  \cite{matsumura}: ``{This simple but important lemma is due to T.\ Nakayama,
  G.\ Azumaya, and W.\ Krull. Priority is obscure, and although it is usually
  called the Lemma of Nakayama, late Prof.\  Nakayama did not like the
  name.''}}
\,\cite{nagata}, Nakayama's Lemma describes one way that a finitely generated
module over an arbitrary commutative ring acts like a vector space over a
field. True to the convention that ``lemma'' often refers to a result serving
as a stepping stone toward another goal, Nakayama's Lemma is applied widely
throughout the field, and the result is typically introduced in a first
graduate course in commutative algebra \cite{atiyah-macdonald, matsumura,
eisenbud}.

@@

\subsection{Constructing the Formal Proof}
Beyond formalizing the relevant structures from higher algebra, we also
formally establish some basic theory from the field in Coq. For instance, 
consider the notion of a \emph{unit} of a
commutative ring $R$, which is an element $a \in R$ with a multiplicative inverse,
i.e., there exists an element $a^{-1} \in R$ for which $x^{-1} \cdot x$ is the
multiplicative identity $1 \in R$.  In fact, it $x$ is a unit, its inverse is unique.

\emily[inline]{Emily: I changed statement to \emph{if and only if} below; do we need to update formal proof?}
Suppose that $I$ is an ideal of a commutative rings $R$.  
We formally state prove the following lemma: If $I$ contains a
unit $a$ if and only if $I$ is the trivial ideal, i.e., $I=R$. 
The informal proof logic is as follows:  Due to the absorption property of $I$,  
$a^{-1}  \cdot a= 1 \in I$.  Hence for every element $r$ of $R$,
$r = r \cdot 1$ is also in $I$, i.e., $I=R$. 
On the other hand, if $I = R$, then the multiplicative identity $1$ is a unit in $I$.

\paragraph{Dealing with the Axiom of Choice.}
Every non-unit element of a commutative ring is contained in some maximal
ideal, a fact that relies on the Axiom of Choice. 
The following standard informal
proof calls upon Zorn's lemma, which is equivalent to the Axiom of Choice, and which 
says that if every chain of elements from a partially ordered set  has an upper
bound in the set, then the set must contain at least one maximal element.

\begin{quote}
	Fix a non-unit $x$ in a commutative ring $R$, and  \(I_{1}\) to be the principal 
	ideal generated by $x$, i.e., the smallest ideal containing the element $x$,
    which consists of all $R$-multiples of it.  No element of $I_1$ is a unit, since if  
    $y \in R$ were the multiplicative inverse of some $R$-multiple
    $r \cdot x$, then $1 = y \cdot (r \cdot x) = (y\cdot r) \cdot x$, showing that $x^{-1} = y \cdot r$, 
    which contradicts the assumption that $x$ is not a unit. Hence by the lemma described earlier in this section, 
    $I_1$ is proper.

    If \(I_{1}\) is not a maximal ideal of $R$, then by definition there exists a strictly larger
    proper ideal \(I_{2}\) of $R$, i.e.,
    $x \in I_{1}\subsetneq I_{2} \subsetneq R$. Likewise, if  \(I_{2}\) is again not
    maximal, then there exists a strictly larger proper ideal
    $I_{3}$, which, if not maximal, is strictly contained in another ideal, $I_4$, etc.

  Continuing in this way, if  this process terminates, we have successfully identified a
    maximal ideal of $R$ that contains $x$, and if not, one obtains an infinite, strictly increasing chain of proper ideals containing $x$, 
    \[I_{1}\subsetneq I_{2}\subsetneq I_{3}\subsetneq\cdots\subsetneq I_n \subsetneq \cdots. \]
    It is straightforward to verify that an increasing union of ideals 
    is, by definition, an ideal containing every ideal in the union.  Moreover, if each ideal in the union
    contains an element $x$, then the union certainly 
       contains \(x\) as well.

      Hence the set $S$ of all ideals of $R$, partially ordered by inclusion, 
       satisfies the property that every chain in $S$ has an upper bound in $S$.
    Therefore Zorn's lemma ensures the existence of a maximal ideal of \(R\) containing \(x\).
\end{quote}
